---
title: "Intro to R"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Intro to R

### Tutorial objectives 

In this tutorial, you will learn:

* what R is
* what R can do 
* some introductory commands 
* what an R packages is 
* some useful R packages 

You can run code in this tutorial directly, or you can choose to run it directly in R on your computer. Your choice!

Ready? Let's go!

### What is R?

R is a **statistical computing language**. See more information (like the history of R) [here](https://www.r-project.org/about.html). In biology, R is used for data analysis, predictions, machine learning, data visualization, etc.

I find that I learn better from *doing* (like doing tutorials), but sometimes you don't know *what to do*! Here are some websites you might want to explore in your free time.

1. <https://education.rstudio.com/learn/beginner/>
2. <https://www.rensvandeschoot.com/tutorials/r-for-beginners/>
3. <https://www.codecademy.com/learn/learn-r> Unfortunately not free :(
4. <https://info.rstudio.com/m000CO20T0IY7SelXxZ0NN0> A talk about how to pick cute side projects to learn R. **Please note: I don't recommend you actually take part in the games at the beginning of the talk...just the R stuff :)**

## What can R do?

R is great with stats, data analysis, and visualization. This means that R is **good at math**. Let's try using R as a calculator!

### R as a calculator

Try out R as a calculator by clicking on "run code".

```{r calc, exercise = TRUE, exercise.eval = FALSE}
## You can use + and - 
2+2
15-7
# you use * to multiply and / to divide
5*4
20/2
# R also has functions for other typical math functions
# squre root
sqrt(182736)
# log2
log2(2.9)
```

### Using variables in R

When you're writing code, you'll want to assign certain numbers as variables for use.

```{r variables, exercise=TRUE}
a <- 2+2
b <- 2*3
a + b 
```

### Test your knowledge

Try writing code where `x=20` and `y=50`. Then subtract `x` from `y`.

```{r variabletest, exercise = TRUE}


```
```{r variabletest-solution}
x <- 20
y <- 50
y-x
```


##  

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

